### 2017-05-08 2017mergingmerges/section/05-Evaluation.tex
Rename Evaluation -> Implementation
Fix section on Fusability after the Great Cutting
benchmark figure

% -----------------------------------------------------------------------------
\eject
\section{Results}
\label{s:Evaluation}

@@

We compare hand-fused against ours, as well as two @Data.Vector@ versions, which uses shortcut fusion.
The shortcut fusion system cannot fuse both operations into a single loop, and both operations require the distances between the line and each point.
% This means a choice must be made: either compute the distances upfront and share them, or recompute the distances in each operation.
% We compared both, and recomputing was significantly faster.
% However, we only benchmarked with two-dimensional points: at higher dimensions, the cost of recomputing distances may outweigh array allocation.

Fig.~\ref{fig:bench:quickhull} shows the runtimes for Quickhull over roughly 80MB of data, or five million points.
The hand-fused version is the fastest, while our version is slower, but still faster than the two vector versions.
The fact that our version is slower than the hand-fused version is surprising, as the generated GHC core is almost identical: the only difference is an extra continuation bound inside the loop, which acts as a sort of indirect jump in the loop body.
% It is possible that recent work on optimising continuations in GHC~\cite{downen2016sequent} will solve this.

% TODO Figure suppressed to work out pagination. 
%      Move all data to a single figure.
% \begin{figure}
% \begin{tikzpicture}
% \begin{axis}[
%   symbolic y coords={Conduit, Pipes, Streaming, Ours, end},
% 	xlabel=Runtime (ms),
%   xmin=0, xmax=290,
% 	enlargelimits=0.01,
% 	xbar interval=0.7,
%   width=7.5cm, height=4.5cm,
% 	legend pos=north west,
% ]
% \addplot coordinates {(282,Conduit) (270,Pipes) (168,Streaming) (86,Ours) (0,end) };
% \end{axis}
% \end{tikzpicture}
% \caption{Runtime for append2; lower is faster.}
% \label{fig:bench:append}
% \end{figure}

The first file benchmark simply appends two files, while counting the number of lines.
We compared against three Haskell streaming libraries: Conduit, Pipes, and Streaming.
Fig.TODO 
% ~\ref{fig:bench:append} 
shows the runtimes for appending 2MB of data.
The absolute performance here is poor because all are using line-buffered IO; in practice one would use chunked IO, but the overhead per chunk would remain.

% TODO: Figure suppressed to work out pagination.
%       Move all data to a single figure.
% \begin{figure}
% \begin{tikzpicture}
% \begin{axis}[
%   symbolic y coords={Streaming, Pipes, Conduit hand, Ours, end},
% 	xlabel=Runtime (ms),
%   xmin=0, xmax=1200,
% 	enlargelimits=0.01,
% 	xbar interval=0.7,
%   width=7.5cm, height=4.5cm,
% 	legend pos=north west,
% ]
% \addplot coordinates {(1200,Streaming) (764,Pipes) (658,Conduit hand) (288,Ours) (0,end) };
% \end{axis}
% \end{tikzpicture}
% \caption{Runtime for partition; lower is faster.}
% \label{fig:bench:part}
% \end{figure}

The second file benchmark takes a file and partitions it into two files: one with even-length lines, and one with odd-length lines.
The output lines are also counted.
The three streaming libraries are pull-based, and do not support multiple outputs: the program must be written in a convoluted way or partially hand-fused.
Even with hand-fusion, the Pipes and Conduit programs are slower than ours, as well as losing any abstraction benefits from using a streaming library.
Fig.TODO 
% ~\ref{fig:bench:part}
shows the runtimes for partitioning a 6MB file.

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
  symbolic y coords={Store, Recompute, Ours, Hand, end},
        xlabel=Runtime (ms),
  xmin=0, xmax=200,
        enlargelimits=0.01,
        xbar interval=0.7,
  width=7.5cm, height=4.5cm,
        legend pos=north west,
]
\addplot coordinates {(200,Store) (130,Recompute) (110,Ours) (75,Hand) (0,end) };
\end{axis}
\end{tikzpicture}
\caption{Runtime for Quickhull; lower is faster.}
\label{fig:bench:quickhull}
\end{figure}

% -----------------------------------------------------------------------------
% \begin{figure}
% \begin{minipage}[t]{0.10\textwidth}
% \end{minipage}
% \begin{minipage}[t]{0.4\textwidth}
% \begin{tikzpicture}
% \begin{axis}[
% % Hide the label on the second graph
% 	ylabel=Process states,
% 	xlabel=Pipelined combinators,
%   ymin=0, ymax=100,
% 	enlargelimits=0.01,
% 	ybar interval=0.7,
%   width=7.5cm, height=4.5cm,
% 	legend pos=north west,
% ]
% \addplot coordinates {(1,9) (2,17) (3,27) (4,37) (5,47) (6,57) (7,67)   (8,1) };
% \addplot coordinates {(1,4) (2,9) (3,14) (4,19) (5,24) (6,29) (7,34)    (8,1) };
% \legend{With merge, Without merge};
% \end{axis}
% \end{tikzpicture}
% \end{minipage}
% \quad
% \quad
% \quad
% \begin{minipage}[t]{0.4\textwidth}
% \begin{tikzpicture}
% \begin{axis}[
% %       ylabel=Process states,
%         xlabel=Parallel combinators,
%   ymin=0, ymax=100,
%         enlargelimits=0.01,
%         ybar interval=0.7,
%   width=7.5cm, height=4.5cm,
%         legend pos=north west,
% ]
% \addplot coordinates {(1,9) (2,27) (3,40) (4,53) (5,66) (6,79) (7,92)
%   % Last bar doesn't show for some reason, so need to add a dummy value for the next one
%     (8,1) };
% 
% \addplot coordinates {(1,4) (2,9) (3,14) (4,19) (5,24) (6,29) (7,34)    (8,1) };
% 
% \legend{With merge, Without merge};
% \end{axis}
% \end{tikzpicture}
% \end{minipage}
% 
% \caption{Maximum output process size for fusing all combinations of up to $n$ combinators.}
% \label{fig:bench:outputsize}
% \end{figure}
% 
% 
% % -----------------------------------------------------------------------------
% \begin{figure}
% 
% \begin{tikzpicture}
% \begin{axis}[
% 	ylabel=Process states,
% 	xlabel=Number of merges,
% %  ymode=log,
%   ymin=0, ymax=1500,
% 	enlargelimits=0.01,
% 	ybar interval=0.7,
%   width=7.5cm, height=4.5cm,
% 	legend pos=north west,
% ]
% % These are the values for splitting.
% % They are smaller than the 'chaining', but look much nicer on the linear graph.
% \addplot coordinates {(1,9) (2,42) (3,97) (4,196) (5,383) (6,746) (7,1461)
%   % (8,2880)
%   (8,1)
%   };
% 
% % These are the values for chaining
% % \addplot coordinates {(1,4) (2,48) (3,194) (4,760) (5,2814) (6,10064) (7,1) };
% 
% \end{axis}
% \end{tikzpicture}
% 
% \caption{Exponential blowup occurs when splitting or chaining merges together.}
% \label{fig:bench:exponential}
% \end{figure}

>>>>

% -----------------------------------------------------------------------------
\section{Implementation}
\label{s:Evaluation}

@@

We compare against a hand-fused program and a @Data.Vector@ program, which uses shortcut fusion.
The shortcut fusion system cannot fuse both operations into a single loop, and both operations recompute the distances between the line and each point.
% This means a choice must be made: either compute the distances upfront and share them, or recompute the distances in each operation.
% We compared both, and recomputing was significantly faster.
% However, we only benchmarked with two-dimensional points: at higher dimensions, the cost of recomputing distances may outweigh array allocation.

The first group in Fig.~\ref{fig:bench:all} shows the runtimes for Quickhull over roughly 80MB of data, or five million points.
The hand-fused version is the fastest, while our version is slower, but still faster than the vector version.
The fact that our version is slower than the hand-fused version is surprising, as the generated GHC core is almost identical: the only difference is an extra continuation bound inside the loop, which acts as a sort of indirect jump in the loop body.
% It is possible that recent work on optimising continuations in GHC~\cite{downen2016sequent} will solve this.

The first file benchmark simply appends two files, while counting the lines.
We compared against three Haskell streaming libraries: Conduit, Pipes, and Streaming.
The second group in Fig.~\ref{fig:bench:all} shows the runtimes for appending 2MB of data.
The absolute performance here is poor because all are using line-buffered IO; in practice one would use chunked IO, but the overhead per chunk would remain.

The second file benchmark takes a file and partitions it into two files: one with even-length lines, and one with odd-length lines.
The output lines are also counted.
The three streaming libraries are pull-based, and do not support multiple outputs: the program must be written in a convoluted way or at least partially hand-fused.
Even with hand-fusion, the Pipes and Conduit programs are slower than ours, as well as losing any abstraction benefits from using a streaming library.
The final group in Fig.~\ref{fig:bench:all} shows the runtimes for partitioning a 1MB file.

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
  symbolic x coords={Quickhull, Append, Partition, end},
        ylabel=Runtime (ms),
  ymin=0, ymax=300,
        enlargelimits=0.01,
        ybar interval=0.7,
  width=7.5cm, height=4.5cm,
  legend style={at={(0.5,-0.2)},anchor=north, legend columns=3}
]
\addplot coordinates { (Quickhull, 75) (Append, 88) (Partition, 60) (end, 0) };
\addplot coordinates {(Quickhull, 110) (Append, 86) (Partition, 61) (end,0)};
\addplot coordinates { (Quickhull, 130) (Append, 0)  };
% \addplot coordinates { (Quickhull, 200) (Append, 0) };
\addplot coordinates {  (Append, 282) (Partition, 136) (end,0) };
\addplot coordinates {  (Append, 270) (Partition, 166) (end,0) };
\addplot coordinates {  (Append, 168) (Partition, 268) (end,0) };
\legend{Handfused, Ours, Vector, Conduit, Pipes, Streaming}
\end{axis}
\end{tikzpicture}
\caption{Runtime for benchmarks; lower is faster.}
\label{fig:bench:all}
\end{figure}

% Quickhull
% \addplot coordinates {(200,Store) (130,Recompute) (110,Ours) (75,Hand) (0,end) };
% Append
% \addplot coordinates {(282,Conduit) (270,Pipes) (168,Streaming) (86,Ours) (0,end) };
% Partition
% \addplot coordinates {(1200,Streaming) (764,Pipes) (658,Conduit hand) (288,Ours) (0,end) };
